R Notebook

### 2017-04-11 doc/RMD/MLG-distribution.Rmd
update MLG-distribution to save SVG files

```{r}
library('tidyverse')
library('assertr')
library('poppr')
library('igraph')
library("ggraph")
library('magrittr')
```

## Loading data and assertions

```{r read_data}
dat <- read.genalex("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", ploidy = 1)
splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
pops <- levels(strata(dat)$Region)
levels(strata(dat)$Region) <- case_when(pops == "FR" ~ "France", 
                                        pops == "MX" ~ "Mexico", 
                                        pops == "AU" ~ "Australia", 
                                        TRUE ~ pops)
dat

repeat_lengths <-
  c(
  `5-2` = 2.000000,
  `5-3` = 0.500000,
  `6-2` = 6.000000,
  `7-2` = 2.000000,
  `8-3` = 2.000000,
  `9-2` = 2.000000,
  `12-2` = 2.000000,
  `17-3` = 3.000000,
  `20-3` = 2.000000,
  `36-4` = 0.500000,
  `50-4` = 0.500000,
  `55-4` = 4.000000,
  `92-4` = 2.000000,
  `106-4` = 4.000000,
  `110-4` = 4.000000,
  `114-4` = 4.000000
  )
repeat_lengths <- ifelse(repeat_lengths < 1, 4, repeat_lengths)
```

```{r excel_data}
ex <- readxl::read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", sheet = "GenAlex", skip = 1) %>%
  select(-1) %>%                # removing first column, which is empty
  gather(locus, allele, -1) %>% # gather all loci into tidy columns
  mutate(locus = trimws(locus)) %>% # remove (F) designator
  mutate(allele = as.integer(allele)) %>% # force alleles to integers
  spread(locus, allele) %>%
  slice(-n())
ex <- ex[!names(ex) %in% locNames(dat)]

# Function to select an adjacent allele. It will select the
# next allele if the next allele is not missing and it's distance
# is one away and the previous allele for the same conditions.
# If none of the conditions are met, it will retain the allele.
cromulent_allele <- Vectorize(function(lower, allele, higher){
  if (!is.na(higher) && abs(allele - higher) == 1){
    out <- higher
  } else if (!is.na(lower) && abs(allele - lower) == 1){
    out <- lower
  } else {
    out <- allele
  }
  out
})
ex
exsummary <- ex %>% 
  gather(locus, allele, -1) %>% # tidy the data
  group_by(locus, allele) %>%   
  summarize(n = n()) %>%        # summarize by count 
  ungroup() %>%
  group_by(locus) %>%           # group the loci, add the lower and upper alleles,
  mutate(lower = lag(allele), higher = lead(allele)) %>% # and then create new_alleles
  mutate(new_allele = ifelse(n < 3, cromulent_allele(lower, allele, higher), allele)) %>%
  select(locus, new_allele, allele)
exsummary
corrected_loci <- ex %>% gather(locus, allele, -1) %>%
  left_join(exsummary, by = c("locus", "allele")) %>%
  mutate(allele = new_allele) %>%
  select(-new_allele) %>%
  spread(locus, allele)
datdf <- genind2df(dat, usepop = FALSE) %>% bind_cols(corrected_loci[-1])
dat  <- df2genind(datdf, ind.names = indNames(dat), strata = strata(dat), ploidy = 1) %>% as.genclone()
```

The incoming strata includes both Severity and Isolate. Since these are not
necessary for delimiting the strata, we will place them in the "other" slot
after converting Severity to numeric. Placing this information in the "other"
slot ensures that these data will travel with the object.

```{r fix_strata}
dat_strata <- strata(dat) %>%
  mutate_all(as.character) %>%
  mutate(Severity = as.numeric(Severity))
strata(dat)     <- select(dat_strata, -Severity, -Isolate)
indNames(dat)   <- dat_strata$Isolate
other(dat)$meta <- select(dat_strata, Severity, Isolate)
```

```{r}
setPop(dat) <- ~Region
dat
keeploci <- !locNames(dat) %in% colnames(corrected_loci)
genotype_curve(dat, sample = 1000, quiet = TRUE)
locus_table(dat)
# rl <- fix_replen(dat, repeat_lengths)
# mlg.filter(dat, distance = bruvo.dist, replen = rl) <- .Machine$double.eps
dat
genotype_curve(dat[loc = keeploci, mlg.reset = TRUE], quiet = TRUE)
# dat <- dat[loc = keeploci, mlg.reset = TRUE]
dat
```

@@

```{r, fig.width = 10, fig.height = 10}
# Plotting the subgraphs
par(mfrow = c(3, 4))
plot_mlg_subgraph(graph16loc[-length(graph16loc)])

@@

@@

```{r, fig.width = 7, fig.height = 7}
par(mfrow = c(1, 1))
(g16o <- cluster_optimal(graph16loc$total))
plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
```

@@

```{r, fig.width = 7, fig.height = 7}
(g11o <- cluster_optimal(graph11loc$total))
plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
```

>>>>

```{r}
library('tidyverse')
library('poppr')
library('igraph')
library("ggraph")
```

## Loading data and setting strata

I'm loading the data and splitting it into the 11 locus and 16 locus versions

```{r}
load("data/sclerotinia_16_loci.dat")
setPop(dat) <- ~Region
dat
locNames(dat)
other(dat)$REPLEN
head(other(dat)$meta)
keeploci <- !locNames(dat) %in% colnames(corrected_loci)
dat11 <- dat[loc = keeploci, mlg.reset = TRUE] # reducing to 11 loci and recalculating mlgs
dat11
```

Now we can inspect the genotype accumulation curves. Notice that the 16 loci
version plateaus, but we know that these extra loci are inconsistent, so we 
might not want to trust them.

```{r genotype_curves, fig.width = 5, fig.height = 5}
locus_table(dat)
genotype_curve(dat, sample = 1000, quiet = TRUE)
genotype_curve(dat11, quiet = TRUE)
```

Creating contracted genotypes for use later

```{r}
mlg.filter(dat, distance = bruvo.dist, replen = other(dat)$REPLEN) <- .Machine$double.eps
dat
mlg.filter(dat11, distance = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps
dat11
mll(dat) <- "original"
mll(dat11) <- "original"
```

@@

```{r subgraphs, fig.width = 10, fig.height = 10}
# Plotting the subgraphs
par(mfrow = c(3, 4))
plot_mlg_subgraph(graph16loc[-length(graph16loc)])

@@

I'm plotting two graphs here because it's useful to see if both of them are
congruent in their community predictions.

@@

```{r graph-16-loci, fig.width = 7, fig.height = 7}
par(mfrow = c(1, 1))
(g16o <- cluster_optimal(graph16loc$total))
plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
ggsave(filename = "results/figures/publication/FigureS1Z.svg", width = 88, units = "mm", scale = 2)
```

@@

```{r graph-11-loci, fig.width = 7, fig.height = 7}
(g11o <- cluster_optimal(graph11loc$total))
plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
ggsave(filename = "results/figures/publication/Figure3Z.svg", width = 88, units = "mm", scale = 2)
```
